Mendelian randomization for causal inference in Genome-wide association studies

### 2022-02-03 mr-review.Rmd
deleted word
added refs to intro

Great interest has developed in inferring causal relationships between complex 
traits in observational human genetics studies. Discovery of such relationships is crucial to 
enhancing our understanding of the biology of health and disease. Two-sample 
Mendelian randomization (MR) studies are a form of instrumental variable analysis in which
genetic markers, typically single nucleotide polymorphisms (SNPs), serve as instruments for 
inferring causal effects of an exposure variable on an outcome variable. They are termed 
"two-sample" MR studies because the exposure variable and the outcome variable are measured on 
two non-overlapping cohorts of subjects. Importantly, the SNPs are measured in both sets of subjects.

Throughout the manuscript, we 
use the term "complex trait" to refer to a trait that doesn't appear to be 
inherited in a Mendelian fashion. Typically, such complex traits, for example, gene expression levels,
are inherited in a polygenic manner, with many genes influencing the trait.

>>>>

Great interest has developed in inferring causal relationships between complex 
traits, *i.e.*, traits that seemingly are not inherited in a Mendelian fashion, in observational human genetics studies. Discovery of such relationships is crucial to 
enhancing our understanding of the biology of health and disease. Two-sample 
Mendelian randomization (MR) studies are a form of instrumental variable analysis in which
genetic markers, typically single nucleotide polymorphisms (SNPs), serve as instruments for 
inferring causal effects of an exposure variable on an outcome variable [@wright1928tariff;@wright1921correlation;@wright1923theory;@stock2003retrospectives;@katan1986apolipoprotein;@katan2004commentary;@didelez2007mendelian]. They are termed 
"two-sample" MR studies because the exposure variable and the outcome variable are measured on 
two non-overlapping cohorts of subjects. Importantly, the SNP instruments are measured in both sets of subjects.

Recent developments have enabled performance of two-sample MR studies with only 
GWAS summary statistics [@burgess2015using].
